On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-06 paper.tex
Submitted
Done!!

@@

\begin{abstract}
We investigate simple oritatami systems in an attempt to establish lower bounds on the size and complexity of computationally universal systems. 
In particular, we look at oritatami systems, where the folding sequence consists of a number of beads of the same type and show that under reasonable assumptions, these systems are not universal.
\end{abstract}

@@

\end{minipage}
\caption{
(Left) RNA origami. 
(Right) an abstraction of its product, i.e., an RNA tile, as a configuration of an oritatami system. A dot $\bullet$ in the figure on the right represents a sequence of 3-4 nucleotides, and the solid arrow and dashed lines represent its RNA transcript and interactions (respectively) based on hydrogen bonds between nucleotides.
}
\label{fig:rna_origami}
\end{figure}

Transcription is the first essential step of gene expression, in which a DNA template sequence is copied into a single stranded RNA sequence by a `molecular Xerox' called RNA polymerase, letter by letter according to the complimentarity relation ${\tt A} \rightarrow {\tt U}$, ${\tt G} \rightarrow {\tt C}$, ${\tt C} \rightarrow {\tt G}$, and ${\tt T} \rightarrow {\tt A}$. 
The copied RNA sequence is called \textit{transcript}. 
The transcript starts folding upon itself into intricate structures immediately after it emerges from the polymerase, as illustrated in Fig.~\ref{fig:rna_origami} (Left).

In a recent breakthrough in molecular engineering by Geary, Rothemund and Andersen~\cite{GearyRothemundAndersen2014} the co-transcriptional folding of RNA is controlled by careful design of the DNA template. As demonstrated in laboratory, this method, called RNA Origami, makes it possible to cotranscriptionally self-assemble a unique RNA rectangular tile highly probably (see Fig.~\ref{fig:rna_origami} (Left)). 
Geary et al.~\cite{GeMeScSe2016} proposed a mathematical model for this process, called oritatami system. 
In this model, an RNA structure is abstracted as a directed path with information on hydrogen-based interaction (bonds) over the triangular grid graph $\mathbb{T}$ as illustrated in Fig.~\ref{fig:rna_origami}. 
An oritatami system folds a transcript of abstract molecules (beads) of finite number of types over $\mathbb{T}$. 
This model has been just proved efficiently Turing universal in \cite{GeMeScSe2018} by simulating cyclic tag systems introduced by Cook~\cite{Cook2004}. 
The simulation involves a very large and complex oritatami system. 
This system is deterministic in the sense that every bead is stabilized uniquely point-wise as well as interaction-wise (for the formal definition, see Sect.~\ref{sec:preliminaries}). 
One future direction of research is to quest for a smaller Turing-universal oritatami system.

@@

    \end{tikzpicture}
    \caption{The zig-zag conformation. This is the only one infinite conformation foldable deterministically by an unary oritatami ssytem at delay 1 and arity 2.}
    \label{TTT_zigzag}
  \end{center}
\end{figure}

@@

An \textit{oritatami system} (OS) $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ is composed of
\begin{itemize}
\item a set $\Sigma$ of bead types, 
\item a rule set $R \subseteq \Sigma \times \Sigma$, 
\item a positive integer $\delta$ called the \textit{delay}, 
\item a positive integer $\alpha$ called the \textit{arity}, 
\item an initial $R$-valid conformation $\sigma \in C_{\le \alpha}(\Sigma)$ called the \textit{seed}, whose first bead is assumed to be at the origin $O$ without loss of generality, 
\item a (possibly infinite) \textit{transcript} $w \in \Sigma^* \cup \Sigma^\omega$, which is to be folded upon the seed by stabilizing beads of $w$ one at a time so as to minimize energy collaboratively with the succeeding $\delta{-}1$ nascent beads. 
\end{itemize}
The energy of a conformation $C = (P, w, H)$, denoted by $\Delta G(C)$, is defined to be ${-}|H|$; the more bonds a conformation has, the more stable it gets. 
The set $\mathcal{F}(\Xi)$ of conformations \textit{foldable} by the system $\Xi$ is recursively defined as: the seed $\sigma$ is in $\mathcal{F}(\Xi)$; and provided that an elongation $C_i$ of $\sigma$ by the prefix $w[1..i]$ be foldable (i.e., $C_0 = \sigma$), its further elongation $C_{i+1}$ by the next bead $w[i+1]$ is foldable if 
\begin{equation}\label{eq:OS_CF}
C_{i+1} \in \argmin_{
\substack{
C \in \mathcal{C}_{\le \alpha} s.t. \\
C_i \xrightarrow{R}_{w[i+1]}C \\
}
}
\min \Bigl\{ \Delta G(C') \Bigm|
C \xrightarrow{R}^*_{w[i+2...i+k]}C', k\le \delta, C' \in \mathcal{C}_{\le \alpha}
\Bigr\}.
\end{equation}
%
Then we say that the bead $w[i+1]$ and the bonds it forms are \textit{stabilized} according to $C_{i+1}$. 
Note that an arity-$\alpha$ oritatami system cannot fold any conformation of arity larger than $\alpha$. 
A conformation foldable by $\Xi$ is \textit{terminal} if none of its elongations is foldable by $\Xi$. 
The oritatami system $\Xi$ is \textit{deterministic} if for all $i \ge 0$, there exists at most one $C_{i+1}$ that satisfies \eqref{eq:OS_CF}. 
A deterministic oritatami system folds into a unique terminal conformation. 
An oritatami system with the empty rule set just folds into an arbitrary elongation of its seed nondeterministically. 
Thus, the rule set is reasonably assumed non-empty.

>>>>

\usepackage{hyperref}

@@

\begin{abstract}
An oritatami system is a novel mathematical model of RNA cotranscriptional folding, which has recently proven extremely significant in information processing in organisms and also controllable artificially in a test tube to construct an artificial structure by folding an RNA sequence.  
This model has turned out to be Turing universal. 
One next step is to simplify the Turing universal oritatami system and another is to characterize weaker oritatami systems as we may not need Turing universality for applications. 
%We investigate simple oritatami systems in an attempt to establish lower bounds on the size and complexity of computationally universal systems. 
In this paper, we look at oritatami systems that folds a unary sequence, and show that under reasonable assumptions, these systems are not universal.
\end{abstract}

@@

\end{minipage}
\caption{
(Left) RNA origami. 
(Right) an abstraction of its product, i.e., an RNA tile, as a configuration of an oritatami system. 
A dot $\bullet$ in the figure on the right represents a sequence of 3-4 nucleotides (oligonucleotides). 
The solid arrow and dashed lines represent respectively its RNA transcript and interactions based on hydrogen bonds between nucleotides.
}
\label{fig:rna_origami}
\end{figure}

Transcription is the first essential step of gene expression, in which a DNA template sequence is copied into a single stranded RNA sequence of nucleotides {\tt A}, {\tt C}, {\tt G}, and {\tt U}  (letter of RNA alphabet) by a `molecular Xerox' called RNA polymerase, nucleotide by nucleotide according to the complimentarity relation ${\tt A} \rightarrow {\tt U}$, ${\tt G} \rightarrow {\tt C}$, ${\tt C} \rightarrow {\tt G}$, and ${\tt T} \rightarrow {\tt A}$. 
The copied RNA sequence is called \textit{transcript}. 
The transcript does NOT remain single-stranded until it is fully synthesized. 
It rather starts folding upon itself into intricate stable conformations (structures) primarily via hydrogen bonds, immediately after it emerges from the polymerase, as illustrated in Fig.~\ref{fig:rna_origami} (Left).

In a recent breakthrough in molecular engineering by Geary, Rothemund and Andersen~\cite{GearyRothemundAndersen2014} the co-transcriptional folding of RNA is controlled by careful design of the DNA template. As demonstrated in laboratory, this method, called RNA Origami, makes it possible to cotranscriptionally self-assemble a unique RNA rectangular tile highly probably (see Fig.~\ref{fig:rna_origami} (Left)). 
This breakthrough and the design of RNA tile has encouraged the research on the nano-scale RNA structure self-assembly so that several successful attempts to self-assemble artificial structures by folding a single-stranded RNA sequence \cite{EKOMSSSO2018,YanYin2017}. 
Geary et al.~\cite{GeMeScSe2016} proposed a mathematical model for this process, called oritatami system. 
In this model, an oligonucleotide (a short sequence of RNA nucleotides) is considered to be a bead and an RNA structure is abstracted as a directed path with information on hydrogen-based interaction (bonds) between beads over the triangular grid graph $\mathbb{T}$ as illustrated in Fig.~\ref{fig:rna_origami}. 
An oritatami system folds a transcript of abstract molecules (beads) of finite number of types over $\mathbb{T}$. 
This model has been just proved efficiently Turing universal in \cite{GeMeScSe2018} by simulating cyclic tag systems introduced by Cook~\cite{Cook2004}. 
The simulation involves a very large and complex oritatami system. 
This system is deterministic in the sense that every bead is stabilized uniquely point-wise as well as interaction-wise (for the formal definition, see Sect.~\ref{sec:preliminaries}). 
One future direction of research is to quest for a smaller Turing-universal oritatami system.

@@

    \end{tikzpicture}
    \caption{The zig-zag conformation. This is the only one infinite conformation foldable deterministically by an unary oritatami system at delay 1 and arity 2.}
    \label{TTT_zigzag}
  \end{center}
\end{figure}

@@

An \textit{oritatami system} (OS) $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ is composed of
\begin{itemize}
\item a set $\Sigma$ of bead types, 
\item a rule set $R \subseteq \Sigma \times \Sigma$, 
\item a positive integer $\delta$ called the \textit{delay}, 
\item a positive integer $\alpha$ called the \textit{arity}, 
\item an initial $R$-valid conformation $\sigma \in C_{\le \alpha}(\Sigma)$ called the \textit{seed}, whose first bead is assumed to be at the origin $O$ without loss of generality, 
\item a (possibly infinite) \textit{transcript} $w \in \Sigma^* \cup \Sigma^\omega$, which is to be folded upon the seed by stabilizing beads of $w$ one at a time so as to minimize energy collaboratively with the succeeding $\delta{-}1$ nascent beads. 
\end{itemize}
The energy of a conformation $C = (P, w, H)$, denoted by $\Delta G(C)$, is defined to be ${-}|H|$; the more bonds a conformation has, the more stable it gets. 
The set $\mathcal{F}(\Xi)$ of conformations \textit{foldable} by the system $\Xi$ is recursively defined as: the seed $\sigma$ is in $\mathcal{F}(\Xi)$; and provided that an elongation $C_i$ of $\sigma$ by the prefix $w[1..i]$ be foldable (i.e., $C_0 = \sigma$), its further elongation $C_{i+1}$ by the next bead $w[i+1]$ is foldable if 
\begin{equation}\label{eq:OS_CF}
C_{i+1} \in \argmin_{
\substack{
C \in \mathcal{C}_{\le \alpha} s.t. \\
C_i \xrightarrow{R}_{w[i+1]}C \\
}
}
\min \Bigl\{ \Delta G(C') \Bigm|
C \xrightarrow{R}^*_{w[i+2...i+k]}C', k\le \delta, C' \in \mathcal{C}_{\le \alpha}
\Bigr\}.
\end{equation}
%
Then we say that the bead $w[i+1]$ and the bonds it forms are \textit{stabilized} according to $C_{i+1}$. 
The easiest way to understand this stabilization process should be the video available at \href{https://www.dailymotion.com/video/x3cdj35}{\tt https://www.dailymotion.com/video/x3cdj35}, in which the Turing universal oritatami system by Geary et al. \cite{GeMeScSe2018}, whose delay is 3, is running. 
Note that an arity-$\alpha$ oritatami system cannot fold any conformation of arity larger than $\alpha$. 
A conformation foldable by $\Xi$ is \textit{terminal} if none of its elongations is foldable by $\Xi$. 
The oritatami system $\Xi$ is \textit{deterministic} if for all $i \ge 0$, there exists at most one $C_{i+1}$ that satisfies \eqref{eq:OS_CF}. 
A deterministic oritatami system folds into a unique terminal conformation. 
An oritatami system with the empty rule set just folds into an arbitrary elongation of its seed nondeterministically. 
Thus, the rule set is reasonably assumed non-empty.
